### 2018-04-27 paper/sat-2017-rebuttals/rebuttal.txt

We thank the reviewers for their detailed comments. Responses are
included below, where each summarized comment is labeled with the
reviews that contained it. Each of the points will be addressed in a
future version of the paper.

--------------------------------------------------------------------

Q2 (Review 1): What is the meaning of "Query_S that is the dependency
of no relation predicate", page 8?

A: An alternative phrasing is that Query_S is not the dependency of
any relational predicate.

---------------------------------------------------------------------

Q3 (Review 1): Does Defnition 6 imply that the dependency graph of
each body is a forest?

A: Not necessarily; the following system is CDD:

(1) R1 => R0
(2) R3 => R1
(3) R4 => R1
(4) R5 => R3
(5) R5 => R4

However, the dependency graph of the body of clause (1) is a DAG.

--------------------------------------------------------------------

Q4 (Review 1): What is Use[R] in "Constructing the Counterexample
Characterization of a CDD"? In this section, why is T used as a
metavariable for an arbitrary CHC system?

A: Use[R] is a boolean predicate to indicate if relational predicate R
occurs in a hypothetical counterexample. The constraints on it are
described in detail in paragraph "Constructing the Counterexample
Characterization of a CDD", page 10.

We have used a fresh variable T so that the counterexample
characterization can be defined once, in general, for an arbitrary
system and instantiated for multple systems, each introduced in
"Constructing the post constraint of a relational predicate."

The variable S is fixed prior to this section as an arbitrary CHC
system to be solved.

--------------------------------------------------------------------

Q5 (Review 1): Can the pre-constraint and post-constraint be
characterized precisely?

A: After giving the intuition behind each constraint, we have given
the formula itself at the end of the paragraph.

---------------------------------------------------------------------

Q1 (Review 1): Was the Expand procedure introduced in previous work?

A: We have included the definition of Expand in Appendix A.

-------------------------------------------------------------------------

Q5 (Review 1): Why can SHARA not parse benchmarks generated by SeaHorn?

A: SHARA correctly parses benchmarks generated by SeaHorn, and these
benchmarks are the subject of our evaluation. The Eldarica CHC solver
[20], a solver to which we attempted to compare SHARA, could not parse
such systems.

---------------------------------------------------------------------------

Q6 (Review 2): Why did you evaluate SHARA compared only to Duality?
Did you try to compare SHARA to other tools?

A: We primarily evaluated SHARA compared to Duality because it is the
most reliable competitive solver that we could find. We also attempted
to compare SHARA to Eldarica on benchmarks generated by SeaHorn;
Eldarica failed to parse such benchmarks. Previous results strongly
indicate that Duality can consistently and significantly outperform
SeaHorn [18]. These were confirmed by our own informal experiments,
which we can report in a future version of this paper.

---------------------------------------------------------------------

Q7 (Review 2): The paper is missing a conclusion and discussion of
potential optimizations as future work.

A: The current paper speculates one potential optimization in Sec. 5,
final paragraph. We will include a more extensive discussion in a
future version of the paper.

---------------------------------------------------------------------

Q8 (Review 3): The paper is in a specialized area of verification that
is unlikely to be interesting to the SAT audience.

A: While we have only discussed one relatively simple example of a
safety verification problem formulated as a CHC system, previous work
has established that CHC systems can formulate an extremely expressive
set of safety-verification problems, including safety of sequential
recursive programs, safety of concurrent programs, and
information-flow security of programs.

CHC solving, which in general is defined as an instance of Constraint
Programming over SMT theories independent of any verification problem,
has also been shown to be useful in modeling a variety of systems,
automata, and concurrent processes [Jaffar], and in knowledge
representation within AI [Brachman].

While the expressive power of CHC systems has been well-studied,
effective techniques for solving them, particularly those based on
interpolation over SMT theories, are relatively new. We believe that
just as techniques communicated between the SAT/SMT and verification
communities for solving SAT and SMT formulas have significantly
strengthened both fields, both communities will benefit from the
mutual study of this branch of Constraint Programming.

---------------------------------------------------------------------

References:

Brachman, Ronald J., Hector J. Levesque, and Raymond Reiter. Knowledge
representation. MIT press, 1992.

Jaffar, Joxan. "A CLP approach to modelling systems." Lecture Notes in
Computer Science (2004).

### 2018-04-27 paper/evaluation.tex

\section{Evaluation}
\label{sec:evaluation}
% describe systems:
We performed an empirical evaluation of \sys to determine how it
compares to existing solvers for recursion-free CHC systems.
%
To do so, we implemented \sys as a modification of \duality CHC
solver, which is included in the \zthree theorem prover~\cite{z3}.
%
We modified \duality to use \sys as its solver for recursion-free CHC
systems.
%
We modified the algorithm used by \duality to generate recursion-free
unwindings of a given recursive system so that in each iteration, it
generates an unwinding with a maximal set of relational predicates and
clauses.
%
In the following context, ``\sys'' refers to \duality modified to
solve general CHC systems.

% describe benchmarks:
We evaluated \sys and \duality on 4,309 CHC systems generated from
programs in the SV-COMP 2015~\cite{svcomp15} verification benchmark
suite.
%
To generate CHC systems, we ran the \seahorn~\cite{gurfinkel15}
verification framework with its default settings (procedures are not
inlined and each loop-free fragment is a clause), set to timeout at 90
seconds.
%
We used the benchmarks in SV-COMP 2015~\cite{svcomp15} because they
were used to evaluate \duality in previous work~\cite{mcmillan14}.

We also attempted to compare \sys to the \eldarica CHC solver, but
\eldarica could not parse the CHC systems generated by \seahorn.
%
We compared \sys and \eldarica on an alternative set of benchmarks
generated by the UFO model checker~\cite{albarghouthi12c}, and found
that \sys outperformed \eldarica by at least an order of magnitude on
an overwhelming number of cases.
%
As a result, we focus our discussion on a comparison of \sys and
\duality.

@@

% analyze data on system size:
\autoref{fig:size} shows the relationship between the solving times of
\duality and \sys and the size of a given system, measured as lines of
code in the format generated by \seahorn.
%
The majority of files have between 1,000 and 100,000 lines, so
\autoref{fig:size} is restricted to this range.
%
The data indicates that that performance improvement of \sys compared
to \duality is consistent across systems of all sizes available.

%result is good
The results indicate that \textbf{(1)} on a significant number of
different verification problems, \sys can perform significantly better than
\duality, but
%
\textbf{(2)} there are some cases in which the strengths of each
algorithm yield better results.
%
We collected the differences between sizes of a given system and its
minimal CDD expansion generated by \sys and found that they were
independent of \sys's performance compared to \duality.
%
Thus, while \duality may in the worst case enumerate exponentially
many derivation trees, it appears to enumerate far fewer
than the worst-case bound in some cases, causing it to perform better
than \sys.
%
Our results indicate that a third approach that combines the strengths
of both \duality and \sys, perhaps by lazily unwinding a given
recursion-free system into a series of CDD systems instead of
derivation trees, could yield further improvements.

>>>>

\section{Evaluation}
\label{sec:evaluation}
% describe systems:
We performed an empirical evaluation of \sys to determine how it
compares to existing CHC solvers.
%
To do so, we implemented \sys as a modification of \duality CHC
solver, which is included in the \zthree theorem prover~\cite{z3}.
%
We modified \duality to use \sys as its solver for recursion-free CHC
systems.
%
We modified the algorithm used by \duality to generate recursion-free
unwindings of a given recursive system so that, in each iteration, it
generates an unwinding which is converted to CDD form.
%
In the following context, ``\sys'' refers to this modified version of
\duality.

% describe benchmarks:
We evaluated \sys and an unmodified version of \duality on 4,309 CHC
systems generated from programs in the SV-COMP 2015~\cite{svcomp15}
verification benchmark suite.
%
To generate CHC systems, we ran the \seahorn~\cite{gurfinkel15}
verification framework with its default settings (procedures are not
inlined and each loop-free fragment is a clause), set to timeout at 90
seconds.
%
We used the benchmarks in SV-COMP 2015~\cite{svcomp15} because they
were used to evaluate \duality in previous work~\cite{mcmillan14}.

We also compared \sys to the \eldarica CHC solver, but
\eldarica could not parse the CHC systems generated by \seahorn.
%
We compared \sys and \eldarica on an alternative set of benchmarks
generated by the UFO model checker~\cite{albarghouthi12c}, and found
that \sys outperformed \eldarica by at least an order of magnitude on
an overwhelming number of cases.
%
As a result, we focus our discussion on a comparison of \sys and
\duality.

@@

% analyze data on system size:
\autoref{fig:size} shows the relationship between the solving times of
\duality and \sys and the size of a given system, measured as lines of
code in the format generated by \seahorn.
%
The majority of files have between 1,000 and 100,000 lines, so
\autoref{fig:size} is restricted to this range.
%
The data indicates that the performance improvement of \sys compared
to \duality is consistent across systems of all sizes available.

%result is good
The results indicate that \textbf{(1)} on a significant number of
different verification problems, \sys can perform significantly better than
\duality, but
%
\textbf{(2)} there are some cases in which the strengths of each
algorithm yield better results.
%
Specifically, the unmodified version of \duality~uses a technique
called lazy annotation to avoid enumerating all derivation trees.
\sys~does not use lazy annotation.
%
We collected the differences between sizes of a given system and its
minimal CDD expansion generated by \sys and found that they were
independent of \sys's performance compared to \duality.
%
Thus, while \duality may in the worst case enumerate exponentially
many derivation trees, it appears to enumerate far fewer
than the worst-case bound in some cases, causing it to perform better
than \sys.
%
Our results indicate that a third approach that combines the strengths
of both \duality and \sys, perhaps by \emph{lazily} unwinding a given
system into a series of CDD systems, could yield further improvements.

### 2018-04-27 paper/introduction.tex

% intro:
\section{Introduction}
\label{sec:intro}
% context, problem defn:
Many critical problems in program verification can be reduced to
solving systems of Constrained Horn Clauses (CHCs), a class of
logic-programming
problems~\cite{bjorner13,flanagan03,rummer13a,rummer13b}.
%
A CHC is a logical implication with the following form:
$$
  R_1(\vec{v_1}) \leftarrow R_2(\vec{v_2}) \land R_3(\vec{v_3}) \land
  ... \land \varphi(\vec{v_0}, \vec{v_1}, \vec{v_2}, \vec{v_3},...)
$$
Here, the left side of the implication, called the head, contains an
uninterpreted relational predicate applied to a vector of variables.
%
The right side has any number of such predicates conjoined together
with a \emph{fact} ($\varphi$). The fact is a logical formula in a background
theory and may use variables named by the predicates.
%
A CHC system is a set of CHCs.
%
The goal of the CHC solving problem is to find suitable interpretations
for each predicate such that each CHC is logically
consistent in isolation.

>>>>

% intro:
\section{Introduction}
\label{sec:intro}
% context, problem defn:
Many critical problems in program verification can be reduced to
solving systems of Constrained Horn Clauses (CHCs), a class of
logic-programming
problems~\cite{bjorner13,flanagan03,rummer13a,rummer13b}.
%
A CHC is a logical implication with the following form:
$$
  R_1(\vec{v_1}) \leftarrow R_2(\vec{v_2}) \land R_3(\vec{v_3}) \land
  ... \land \varphi(\vec{v_0}, \vec{v_1}, \vec{v_2}, \vec{v_3},...)
$$
Here, the left side of the implication, called the head, contains an
uninterpreted relational predicate applied to a vector of variables.
%
The right side has any number of such predicates conjoined together
with a \emph{constraint} ($\varphi$). The constraint is a logical formula in a background
theory and may use variables named by the predicates.
%
A CHC system is a set of CHCs.
%
The goal of the CHC solving problem is to find suitable interpretations
for each predicate such that each CHC is logically
consistent in isolation.
